\section{Definitions and Standards}

\subsection{Color Palette Structure}

A palette $P$ comprises $n$ monochromatic families:
\[
P = \{ S_1, S_2, \dots, S_n \}.
\]

Each color family $S_j$ is a \emph{monochromatic ramp} consisting of $N$ lightness steps, generated from a single base color. Formally,
\[
S_j = \{ c_{j,0}, c_{j,1}, \dots, c_{j,N-1} \}.
\]

Colors within each family are arranged in a strictly monotonic order of perceived lightness, forming a one-dimensional lightness ramp. Within $S_j$, there exists a unique base color $C_{\text{base},j} \in S_j$ satisfying
\[
C_{\text{base},j} =
\arg\max_{c_{j,i} \in S_j} C^*(c_{j,i}).
\]

This color possesses the maximum chroma in the range and serves as the representative hue for the entire family. By definition, $S_j$ is assumed to be the result of a fixed-hue interpolation where only lightness and its dependent chroma vary. This assumption ensures consistent reference quantities
($L_{\text{base}}, h_{\text{base}}, C_{\text{base}}$)
for all subsequent metrics.

\subsection{WCAG Contrast Ratio}

According to WCAG~2.1~\cite{wcag21}, the contrast ratio is defined based on relative luminance $Y \in [0,1]$:
\[
CR(c_1, c_2) =
\frac{Y_1 + 0.05}{Y_2 + 0.05},
\quad Y_1 \ge Y_2.
\]

Relative luminance is calculated as
\[
Y =
0.2126\,R_{\text{linear}}
+ 0.7152\,G_{\text{linear}}
+ 0.0722\,B_{\text{linear}}.
\]

Standard contrast ratio thresholds include:
\begin{itemize}
  \item \textbf{3:1} for UI components and large text.
  \item \textbf{4.5:1} for body text (AA standard).
  \item \textbf{7:1} for enhanced readability (AAA).
\end{itemize}

This framework identifies the \textbf{4.5:1 threshold} as the primary benchmark for comparative evaluation.

\subsection{Contrast Span}

The \emph{Contrast Span} $K$ of a hue family is the minimum index distance that guarantees a contrast ratio $\ge 4.5{:}1$ for all color pairs:
\[
K =
\min \left\{
k \in \mathbb{N} :
CR(c_i, c_{i+k}) \ge 4.5,
\;\forall i \in [0, N-k-1]
\right\}.
\]

For multi-family palettes, the global span is defined as
\[
K_{\text{palette}} = \max_{j=1}^{n} K_j.
\]

This metric enables predictable accessibility: designers can ensure compliant contrast by selecting colors separated by at least $K$ steps without requiring runtime calculations.
As illustrated in Table~\ref{tab:contrast-span}, a fixed span guarantees WCAG-compliant contrast across all valid index pairs.

\input{tables/contrast-span}
