\subsection{Hue Stability ($\mathcal{H}$)}

The \textbf{Hue Stability} metric evaluates the consistency of hue in a monochromatic
color scale as lightness varies. A palette is considered stable if hue fluctuates
minimally around a reference color rather than drifting toward other hues.

\subsubsection{Hue Representation and Periodicity Handling}

Hue in CIELAB space is determined by
\begin{equation}
h_i = \mathrm{atan2}(b^*_i, a^*_i).
\end{equation}

Since hue is a periodic angular quantity, the hue sequence is \textbf{unwrapped}
to remove discontinuities at $360^\circ$ and ensure continuity:
\begin{equation}
h^{\mathrm{unwrap}}_i =
h^{\mathrm{unwrap}}_{i-1} + \Delta h_i,
\qquad
\Delta h_i \in (-180^\circ, 180^\circ].
\end{equation}

This unwrapping ensures that hue deviations reflect actual shifts rather than
artifacts from angular periodicity.

\subsubsection{Endpoint Exclusion}

Since the first and last steps of a design-system ramp are typically achromatic
anchors (pure white and pure black), their hue angles are numerically undefined
or arbitrary ($C^* \approx 0$). Including them would inject noise unrelated to
the palette's chromatic identity. Accordingly, the hue stability metric is
evaluated only over the \emph{inner} steps $\{c_1, \dots, c_{N-2}\}$, yielding
$N' = N - 2$ observations.

\subsubsection{Reference Color and Hue Deviation}

The reference hue $h_{\mathrm{base}}$ is chosen as the hue of the
\textbf{color with maximum chroma} in the palette, as this typically represents
the dominant hue identity.

The angular distance between each step and the reference color is defined as
\begin{equation}
d_i =
\min\!\left(
|h_i - h_{\mathrm{base}}|,
\; 360^\circ - |h_i - h_{\mathrm{base}}|
\right),
\end{equation}
where $d_i \in [0, 180^\circ]$ represents the hue deviation at step $i$.

\subsubsection{Normalization Envelope}

To normalize deviation magnitudes into a $[0,1]$ score, a fixed reference
envelope is constructed representing systematic linear drift from zero
deviation at one end of the ramp to the maximum distinguishable angular
distance ($180^\circ$) at the other:
\begin{equation}
d^{\max}_i = 180^\circ \cdot \frac{i}{N' - 1}.
\end{equation}

This envelope serves as a \textbf{normalization scale} rather than a strict
theoretical maximum: its purpose is to provide a consistent denominator that
is independent of palette content, ensuring that scores remain comparable
across palettes with different step counts and base-color positions.
Because the envelope is monotonically increasing, palettes whose base color
lies near the ramp center will have small actual deviations $d_i$ relative
to $d^{\max}_i$ on both sides of the base, yielding high stability scores
as intended.

\subsubsection{Hue Stability Metric Definition}

The root mean square error of hue deviation is computed over the $N'$ inner steps:
\begin{equation}
RMSE =
\sqrt{
\frac{1}{N'}
\sum_{i=0}^{N'-1} d_i^2
}.
\end{equation}

The corresponding worst-case error is
\begin{equation}
RMSE_{\max} =
\sqrt{
\frac{1}{N'}
\sum_{i=0}^{N'-1} \left(d^{\max}_i\right)^2
}.
\end{equation}

The \textbf{Hue Stability} metric is defined by
\begin{equation}
\mathcal{H} =
\max\!\left(
0,\;
1 - \frac{RMSE}{RMSE_{\max}}
\right).
\end{equation}

\subsubsection{Interpretation}

\begin{itemize}
  \item $\mathcal{H} \approx 1$: Hue remains stable around the reference color
  throughout the palette.
  \item $\mathcal{H} \to 0$: Strong hue drift approaching the worst-case linear
  scenario.
\end{itemize}

This metric is independent of lightness and chroma, measuring only the
\textbf{geometric stability of hue} in color space.
