\section{Conclusion}

This work introduces a unified quantitative framework for evaluating the
quality of sequential monochromatic color palettes, grounded in the CIELAB
color space and motivated by both perceptual uniformity and accessibility
requirements. Five complementary metrics are proposed, covering contrast
efficiency under WCAG~4.5:1 constraints, lightness linearity, chroma
smoothness, hue stability, and perceptual spacing uniformity. Each metric
captures a distinct structural or perceptual property and is normalized to
enable consistent aggregation.

Through large-scale benchmarking of widely used industry design systems, the
framework reveals systematic differences in how palettes allocate lightness
range, distribute perceptual steps, and preserve hue identity. In particular,
the analysis identifies contrast span $K$ as a previously under-formalized
but critical parameter, with empirical results indicating that spans close to
$(N-1)/2$ maximize accessibility efficiency while preserving usable color
density. These findings demonstrate that many high-quality palettes converge
toward similar structural ratios, despite differing design origins.

The proposed composite score, based on geometric aggregation, enables
objective comparison without allowing compensation between weak and strong
dimensions, thereby reflecting overall palette completeness rather than
isolated excellence. As a result, the framework supports reproducible
benchmarking, automated palette validation, and optimization-driven color
system design.

\subsection{Limitations}

Several limitations should be acknowledged.
First, while CIELAB is widely adopted as a standard perceptual color space,
it exhibits known non-uniformities---particularly in the blue
region~\cite{luo2001}---that may affect chroma and hue measurements for
certain hue families. More recent color appearance models such as
CAM16 or perceptually uniform spaces like Oklab~\cite{safdar2017} may offer
improved uniformity, though at the cost of reduced compatibility with the
CIEDE2000 metric.
Second, the framework evaluates only \emph{sequential monochromatic} ramps
and does not address diverging, categorical, or multi-hue palettes, which
require fundamentally different structural assumptions.
Third, the five proposed metrics carry equal weight in the geometric mean
composite score; domain-specific applications may benefit from weighted
aggregation, though determining appropriate weights would require empirical
user studies beyond the scope of this work.
Finally, the benchmark evaluation is limited to design systems that publish
algorithmically constructed sequential ramps; systems relying on hand-picked
semantic tokens were excluded, which may limit generalizability.

\subsection{Future Work}

Future work will extend the framework to diverging and
multivariate palettes, and investigate how analogous metric formulations may
be adapted to alternative perceptually uniform color spaces (e.g., CAM02-UCS
or CAM16) while preserving cross-space comparability.
Empirical validation through user studies comparing metric rankings with
human preference judgments would further strengthen the framework's
applicability.
