\begin{thebibliography}{10}

\bibitem{iso11664}
ISO/CIE.
\newblock \emph{Colorimetry — Part 4: CIE 1976 L*a*b* Colour Space}.
\newblock ISO/CIE 11664-4:2019, 2019.

\bibitem{fritsch1980}
F.~N. Fritsch and R.~E. Carlson.
\newblock Monotone piecewise cubic interpolation.
\newblock \emph{SIAM Journal on Numerical Analysis}, 17(2):238--246, 1980.

\bibitem{high2023}
R.~High, P.~Green, and P.~Nussbaum.
\newblock The Helmholtz--Kohlrausch effect on display-based light colors and
simulated substrate colors.
\newblock \emph{Color Research \& Application}, 48(3):341--354, 2023.

\bibitem{zeileis2009}
A.~Zeileis, K.~Hornik, and P.~Murrell.
\newblock Escaping RGBland: Selecting colors for statistical graphics.
\newblock \emph{Computational Statistics \& Data Analysis}, 53(9):3259--3270,
2009.

\bibitem{luo2001}
M.~R. Luo, G.~Cui, and B.~Rigg.
\newblock The development of the CIE 2000 colour-difference formula: CIEDE2000.
\newblock \emph{Color Research \& Application}, 26(5):340--350, 2001.

\bibitem{wcag21}
W3C.
\newblock \emph{Web Content Accessibility Guidelines (WCAG) 2.1}.
\newblock 2018.
\newblock Available at \url{https://www.w3.org/TR/WCAG21/}.

\bibitem{safdar2017}
M.~Safdar, G.~Cui, Y.~J. Kim, and M.~R. Luo.
\newblock Perceptually uniform color space for image signals including high
dynamic range and wide gamut.
\newblock \emph{Optics Express}, 25(13):15131--15151, 2017.

\bibitem{mach1865}
E.~Mach.
\newblock Über die Wirkung der räumlichen Verteilung des Lichtreizes auf die
Netzhaut.
\newblock \emph{Sitzungsberichte der Kaiserlichen Akademie der Wissenschaften,
Wien}, 52:303--322, 1865.

\bibitem{epperson1987}
J.~F. Epperson.
\newblock On the Runge example.
\newblock \emph{The American Mathematical Monthly}, 94(4):329--341, 1987.

\bibitem{srgb}
IEC.
\newblock \emph{Multimedia systems and equipment — Colour measurement and
management — Part 2-1: Colour management — Default RGB colour space — sRGB}.
\newblock IEC 61966-2-1:1999, 1999.

\bibitem{gramazio2017}
C.~C. Gramazio, D.~H. Laidlaw, and K.~B. Schloss.
\newblock Colorgorical: Creating discriminable and preferable color palettes for
information visualization.
\newblock \emph{IEEE Transactions on Visualization and Computer Graphics},
23(1):521--530, 2017.

\bibitem{yang2024}
F.~Yang, W.~Zhang, and X.~Chen.
\newblock Computational color theme scoring via user preference modelling.
\newblock \emph{ACM Transactions on Applied Perception}, 21(2), 2024.

\bibitem{norman2024}
J.~Norman, A.~Klokmose, and B.~Victor.
\newblock Purpose, palette, prototype: Understanding professional color design
workflows.
\newblock In \emph{Proceedings of the ACM CHI Conference on Human Factors in
Computing Systems}, 2024.

\bibitem{colorbuilder2019}
S.~Lin and J.~Fortuna and C.~Kulkarni and M.~Stone and J.~Heer.
\newblock Modeling Color Difference for Visualization Design.
\newblock In \emph{Proceedings of the ACM CHI Conference on Human Factors in
Computing Systems}, 2019.

\bibitem{stone2014}
M.~Stone, D.~Albers Szafir, and V.~Setlur.
\newblock An engineering model for color difference as a function of size.
\newblock In \emph{Proceedings of the 22nd Color and Imaging Conference},
pages 253--258, 2014.

\end{thebibliography}
